### 2021-08-30 paper/localpsf.tex
some reorganization of intro

for arbitrary functions $u$ and $v$. After discretization (see Appendix \ref{app:discretized_operations}), $\Aop$ becomes a dense matrix, $\mathbf{A}$, that is typically too large to be built and stored. This property is called ``matrix-free'' because at the discrete level it means that one can perform matrix-vector products of $\mathbf{A}$ and $\mathbf{A}^T$ with arbitrary vectors, but one cannot easily access matrix entries $\mathbf{A}_{ij}$. One could compute $\mathbf{v} = \mathbf{A}\mathbf{e}_j$, where $\mathbf{e}_j=(0,\dots,0,1,0,\dots,0)$ is the unit vector which has $j^\text{th}$ component equal to one and all other components equal to zero, then extract $\mathbf{A}_{ij} = \mathbf{v}_i$. However, this process is wasteful because one computes the entire vector $\mathbf{v}$, then discards all but one component. In applications, applying $\mathbf{A}$ to a vector may involve iteratively solving a large linear system, or timestepping, or performing some other nontrivial computational procedure. For example, in reduced space approaches to PDE-constrained optimization and inverse problems, Hessian information is available only through the application of the Hessian to vectors, and one such Hessian application requires solving two PDEs \cite{HESSIANADJOINT}. 

@@

In this paper we present a computationally efficient method to form product-convolution approximations of operators possessing these four properties. Once the product-convolution approximation is formed, we convert it to hierarchical matrix (H-matrix) format, and use fast H-matrix methods to perform further useful linear algebra operations.

@@

The most popular existing Hessian approximation methods are based on forming a low rank approximation of the data misfit term in the Hessian, or the data misfit term preconditioned by the prior term \cite{CCGOPAPERS}. Either the Lanczos method or the randomized singular value decomposition \cite{HMTRANDOM} are used to perform the low rank approximation using only matrix-vector products. These methods suffer from a ``data predicament''---if the data are highly informative about the unknown parameter, then the numerical rank of the data misfit term in the Hessian is large, so a large number of matrix-vector products are required to form the aforementioned low rank approximation. The ideal scenario from a scientific perspective (highly informative data) is the worst case scenario from a computational perspective (large computational cost) \cite{MYDISSERTATION}. Although the method we present in this paper is not applicable to all Hessians, it is applicable to several Hessians of practical interest. For these Hessians, our method offers a \emph{data-scalable} alternative to conventional low-rank Hessian approximation methods, because our method can form high-rank approximations of an operator using a small number of matrix-vector products.

\subsection{Product-convolution approximation}

>>>>

for arbitrary functions $u$ and $v$. After discretization (see Appendix \ref{app:discretized_operations}), $\Aop$ becomes a dense matrix, $\mathbf{A}$, that is typically too large to be built and stored. This property is called ``matrix-free'' because at the discrete level it means that one can perform matrix-vector products of $\mathbf{A}$ and $\mathbf{A}^T$ with arbitrary vectors, but one cannot easily access matrix entries $\mathbf{A}_{ij}$. One could compute $\mathbf{v} = \mathbf{A}\mathbf{e}_j$, where $\mathbf{e}_j=(0,\dots,0,1,0,\dots,0)$ is the unit vector which has $j^\text{th}$ component equal to one and all other components equal to zero, then extract $\mathbf{A}_{ij} = \mathbf{v}_i$. However, this process is wasteful because one computes the entire vector $\mathbf{v}$, then discards all but one component. In applications, applying $\mathbf{A}$ to a vector may involve iteratively solving a large linear system, timestepping, or performing some other nontrivial computational procedure. For example, in reduced space approaches to PDE-constrained optimization and inverse problems, Hessian information is available only through the application of the Hessian to vectors, and one such Hessian application requires solving two PDEs \cite{HESSIANADJOINT}. 

@@

Our overall strategy is to form a product-convolution approximation, $\AopPc$, of $\Aop$ using only operator actions, then convert $\AopPc$ into an H-matrix, which we denote by $\AmatPc$. Schematically,
\begin{equation*}
\Aop \longrightarrow \underbrace{\AopPc}_{\substack{\text{product}\\ \text{convolution}}} \longrightarrow \underbrace{\AmatPc}_{\text{H-matrix}}.
\end{equation*}
We will define product-convolution approximations and H-matrices, and discuss them in greater detail, throughout the paper. The novel contributions of our method are in the way that we form the product-convolution approximation ($\Aop \rightarrow \AopPc$). This is the part of the method where we use the four properties of $\Aop$ described above. Fast conversion of $\AopPc$ to H-matrix format ($\AopPc \rightarrow \AmatPc$) is possible because integral kernel entries for $\AopPc$ are easily accessible (unlike integral kernel entries for $\Aop$). Once in H-matrix format, we use fast H-matrix methods to perform further linear algebra operations.

@@

The most popular existing Hessian approximation methods are based on forming a low rank approximation of the data misfit term in the Hessian, or the data misfit term preconditioned by the prior term \cite{CCGOPAPERS}. Either the Lanczos method or the randomized singular value decomposition \cite{HMTRANDOM} are used to perform the low rank approximation using only matrix-vector products. These methods suffer from a ``data predicament''---if the data are highly informative about the unknown parameter, then the numerical rank of the data misfit term in the Hessian is large, so a large number of matrix-vector products are required to form the aforementioned low rank approximation. The ideal scenario from a scientific perspective (highly informative data) is therefore the worst case scenario from a computational perspective (large computational cost) \cite{MYDISSERTATION}. Although the method we present in this paper is not applicable to all Hessians, it is applicable to several Hessians of practical interest. For these Hessians, our method offers a \emph{data-scalable} alternative to conventional low-rank Hessian approximation methods, because our method can form high-rank approximations of an operator using a small number of matrix-vector products.

\subsection{Product-convolution approximation}
\label{sec:prodconv_intro}
